---
title: "R Notebook"
output: html_notebook
---
# Libraries
```{r}
library(tidyverse)
library(janitor)
library(missRanger)
library(ranger)
library(dlookr)
library(pROC)
library(glmulti)
library(modelr)
library(randomForest)
```
#Data
```{r}
aggregate_dataset <- read_csv("data/shs_aggregate_responses.csv") %>% clean_names()
```


## Notes about the dataset
- There are some NAs in the data, namely in the following rows
    - Satisfaction with nearest greenspace (9988 missing)
    - Economic status (1 missing)
    - Highest education level (10920 missing)
    - Nearest green space use (16766 missing)
    - Volunteering last twelve months (6778)
    
Can we meaningfully impute these?

- 
```{r}
plot_na_pareto(aggregate_dataset)
```
```{r}
plot_na_intersect(aggregate_dataset)
```




Community belonging - Don't know - 576
Neighbourhood rating - No opinion - 190
Satisfaction with nearest greenspace - No opinion - 3418
Nearest green space use - Don't know - 83
Distance to nearest green space - Don't know - 386


To Impute or not Impute: That’s the Question - Lodder, 2013
"If more than
25% of the data is missing and researchers apply modern
treatments to impute the missing data, then they should
always compare the results of their subsequent analyses
with the results they would have obtained if they had
used complete case analysis." (Lodder, 2013)


So we can try using the MissRanger package to compute missing categorical variables using the random forest method (https://yuzar-blog.netlify.app/posts/2021-03-04-how-to-impute-missing-values-in-r/#multivariate-imputation-of-categorical-variables)

We can then compare the results to the complete case analysis (only complete observations are included - meaning all NAs are dropped)

## Test/train split

```{r}
n_data <- nrow(aggregate_dataset)
test_index <- sample(1:n_data, size = n_data*0.2)
agg_dataset_test <- slice(aggregate_dataset, test_index)
agg_dataset_train <- slice(aggregate_dataset, -test_index)
```

### Imputation for train dataset
```{r}
train_dataset_imputed <- missRanger(
  agg_dataset_train,
  formula = . ~ .,
  num.trees = 100,
  verbose = 2,
  seed = 111,
  returnOOB = TRUE
  
)
```
### Imputation for test dataset
```{r}
test_dataset_imputed <- missRanger(
  agg_dataset_test,
  formula = . ~ .,
  num.trees = 100,
  verbose = 2,
  seed = 111,
  returnOOB = TRUE
)
```

# Community belonging


```{r}
train_comm_belonging <- train_dataset_imputed %>% 
  #take out don't know b/c we want to predict strong/weak belonging
  filter(community_belonging != "Don't know") %>% 
   #create binary version of community belonging
  mutate(community_belonging_binary = if_else(community_belonging %in% c(
    "Very strongly", "Fairly strongly"), "Strong", "Not strong"), .after = 
      community_belonging) %>%
  #factor all the categorical variables
  mutate(neighbourhood_rating = factor(neighbourhood_rating, levels = 
                                         c("Very good",
                                           "Fairly good",
                                           "Fairly poor",
                                           "Very poor",
                                           "No opinion")),
         distance_to_nearest_green_space = factor(distance_to_nearest_green_space,
                                                  levels = c(
                                                    "A 5 minute walk or less",
                                                    "Within a 6-10 minute walk",
                                                    "Within an 11-20 minute walk",
                                                    "Within a 21-30 minute walk",
                                                    "More than a 30 minute walk away",
                                                    "Don't know"
                                                  )),
         satisfaction_with_nearest_green_space = factor(
           satisfaction_with_nearest_green_space,
           levels = c("Very satisfied",
                      "Fairly satisfied",
                      "Neither satisfied nor dissatisfied",
                      "Fairly dissatisfied",
                      "Very dissatisfied",
                      "No opinion"
                      )),
         age = factor(age, levels = c(
           "16 - 34 Years",
           "35 - 64 Years",
           "65 + Years"
         )),
         gender = as.factor(gender),
         economic_status = as.factor(economic_status),
         highest_education_level = factor(highest_education_level, levels = 
                                            c(
                                              "Standard grade or equiv (SVQ level 1 or 2).",
                                              "Higher, A level or equivalent (SVQ Level 3)",
                                              "HNC/HND or equivalent (SVQ Level 4)",
                                              "Degree, Professional qualification (Above SVQ Level 4)",
                                              "Other qualification"
                                            )),
         nearest_green_space_use = as.factor(nearest_green_space_use),
         volunteering_last_twelve_months = as.factor(volunteering_last_twelve_months),
         community_belonging_binary = as.factor(community_belonging_binary)
         ) %>% 
  #take out original community belonging and the n_persons column (the number of people with the same observed characteristics isn't useful to the model)
  select(-c(community_belonging, n_persons))
```


```{r}
community_rf_classifier <- ranger(community_belonging_binary ~ .,
                                  data = train_comm_belonging,
                                  importance = "impurity",
                                  num.trees = 500)
```

```{r}
community_rf_classifier
```

```{r}
importance(community_rf_classifier)
```

```{r}
community_rf_classifier$confusion.matrix
```

```{r}
#testing to see whether a different value for mtry will be better than the default
oob_values <- vector(length = 10)
for(i in 1:10) {
  temp.model <- ranger(community_belonging_binary ~ .,
                       data = train_comm_belonging,
                       importance = "impurity",
                       num.trees = 500,
                       mtry = i)
  oob_values[i] <- temp.model$prediction.error
}
oob_values
```

```{r}
community_rf_classifier <- ranger(community_belonging_binary ~ .,
                                  data = train_comm_belonging,
                                  importance = "permutation",
                                  num.trees = 50,
                                  mtry = 2)
```

```{r}
importance(community_rf_classifier)
```


```{r}
head(importance_pvalues(community_rf_classifier, formula = community_belonging_binary ~ .,
                        data = train_comm_belonging, method = "altmann"))
```




# Neighbourhood rating

## Prepare training dataset
```{r}
train_neighbourhood_rating <- train_dataset_imputed %>% 
  #take out don't know b/c we want to predict good/poor ratings
  filter(neighbourhood_rating != "No opinion") %>% 
   #create binary version of neighbourhood rating
  mutate(neighbourhood_rating_binary = if_else(neighbourhood_rating %in% c(
    "Very good", "Fairly good"), "Good", "Poor"), .after = 
      neighbourhood_rating) %>%
  #factor all the categorical variables
  mutate(community_belonging = factor(community_belonging, levels = 
                                         c("Very strongly",
                                           "Fairly strongly",
                                           "Not very strongly",
                                           "Not at all strongly",
                                           "Don't know")),
         distance_to_nearest_green_space = factor(distance_to_nearest_green_space,
                                                  levels = c(
                                                    "A 5 minute walk or less",
                                                    "Within a 6-10 minute walk",
                                                    "Within an 11-20 minute walk",
                                                    "Within a 21-30 minute walk",
                                                    "More than a 30 minute walk away",
                                                    "Don't know"
                                                  )),
         satisfaction_with_nearest_green_space = factor(
           satisfaction_with_nearest_green_space,
           levels = c("Very satisfied",
                      "Fairly satisfied",
                      "Neither satisfied nor dissatisfied",
                      "Fairly dissatisfied",
                      "Very dissatisfied",
                      "No opinion"
                      )),
         age = factor(age, levels = c(
           "16 - 34 Years",
           "35 - 64 Years",
           "65 + Years"
         )),
         gender = as.factor(gender),
         economic_status = as.factor(economic_status),
         highest_education_level = factor(highest_education_level, levels = 
                                            c(
                                              "Standard grade or equiv (SVQ level 1 or 2).",
                                              "Higher, A level or equivalent (SVQ Level 3)",
                                              "HNC/HND or equivalent (SVQ Level 4)",
                                              "Degree, Professional qualification (Above SVQ Level 4)",
                                              "Other qualification"
                                            )),
         nearest_green_space_use = as.factor(nearest_green_space_use),
         volunteering_last_twelve_months = as.factor(volunteering_last_twelve_months),
         neighbourhood_rating_binary = as.factor(neighbourhood_rating_binary)
         ) %>% 
  #take out original neighbourhood rating and the n_persons column (the number of people with the same observed characteristics isn't useful to the model)
  select(-c(neighbourhood_rating, n_persons))
```

## Random forest model



```{r}
#testing to see which value for mtry will produce the lowest error
oob_values <- vector(length = 7)
for(i in 1:7) {
  temp.model <- ranger(neighbourhood_rating_binary ~ .,
                       data = train_neighbourhood_rating,
                       importance = "impurity",
                       num.trees = 500,
                       mtry = i)
  oob_values[i] <- temp.model$prediction.error
}
oob_values
```
Based on this output, a value of mtry = 2 is the best for this data.

```{r}
#grow the random forest
neighbourhood_rf_classifier <- ranger(neighbourhood_rating_binary ~ .,
                                  data = train_neighbourhood_rating,
                                  importance = "impurity",
                                  num.trees = 1000,
                                  mtry = 2)
```

```{r}
#get importance values
ranger::importance(neighbourhood_rf_classifier)
```

```{r}
#get p values for importance
neighbourhood_rf_100 <- ranger(neighbourhood_rating_binary ~ .,
                                  data = train_neighbourhood_rating,
                                  importance = "permutation",
                                  num.trees = 100,
                                  mtry = 2)

importance_pvalues(neighbourhood_rf_100, 
                   formula = neighbourhood_rating_binary ~ ., 
                   method = "altmann",
                   data = train_neighbourhood_rating,
                   num.permutations = 50)
```
It looks like the most important variables were year, community belonging, and satisfaction with nearest greenspace - all the rest were nonsignificant

```{r}
neighbourhood_rf_classifier$confusion.matrix
```
```{r}
#true positive rate
43588/(43588 + 73)
#true negative rate
167/(167 + 2845)
#false positive rate
2845/(167 + 2845)
#false negative rate
73/(73 + 43588)
```
```{r}
#get random forest model from random forest package using same formula, this will allow us to get probability predictions
rf_nrating_classifier <- randomForest(formula = neighbourhood_rating_binary ~ .,
                                      data = train_neighbourhood_rating,
                                      ntree = 1000,
                                      norm.votes = FALSE)
```

```{r}
rf_nrating_classifier
randomForest::varImpPlot(rf_nrating_classifier)
```
We're getting slightly different variables that are most important in the model generated with random forest package

### Test on the test dataset
```{r}
#Modify the test dataset using the same method as before
test_neighbourhood_rating <- test_dataset_imputed %>% 
  #take out don't know b/c we want to predict good/poor ratings
  filter(neighbourhood_rating != "No opinion") %>% 
   #create binary version of neighbourhood rating
  mutate(neighbourhood_rating_binary = if_else(neighbourhood_rating %in% c(
    "Very good", "Fairly good"), "Good", "Poor"), .after = 
      neighbourhood_rating) %>%
  #factor all the categorical variables
  mutate(community_belonging = factor(community_belonging, levels = 
                                         c("Very strongly",
                                           "Fairly strongly",
                                           "Not very strongly",
                                           "Not at all strongly",
                                           "Don't know")),
         distance_to_nearest_green_space = factor(distance_to_nearest_green_space,
                                                  levels = c(
                                                    "A 5 minute walk or less",
                                                    "Within a 6-10 minute walk",
                                                    "Within an 11-20 minute walk",
                                                    "Within a 21-30 minute walk",
                                                    "More than a 30 minute walk away",
                                                    "Don't know"
                                                  )),
         satisfaction_with_nearest_green_space = factor(
           satisfaction_with_nearest_green_space,
           levels = c("Very satisfied",
                      "Fairly satisfied",
                      "Neither satisfied nor dissatisfied",
                      "Fairly dissatisfied",
                      "Very dissatisfied",
                      "No opinion"
                      )),
         age = factor(age, levels = c(
           "16 - 34 Years",
           "35 - 64 Years",
           "65 + Years"
         )),
         gender = as.factor(gender),
         economic_status = as.factor(economic_status),
         highest_education_level = factor(highest_education_level, levels = 
                                            c(
                                              "Standard grade or equiv (SVQ level 1 or 2).",
                                              "Higher, A level or equivalent (SVQ Level 3)",
                                              "HNC/HND or equivalent (SVQ Level 4)",
                                              "Degree, Professional qualification (Above SVQ Level 4)",
                                              "Other qualification"
                                            )),
         nearest_green_space_use = as.factor(nearest_green_space_use),
         volunteering_last_twelve_months = as.factor(volunteering_last_twelve_months),
         neighbourhood_rating_binary = as.factor(neighbourhood_rating_binary)
         ) %>% 
  #take out original neighbourhood rating and the n_persons column (the number of people with the same observed characteristics isn't useful to the model)
  select(-c(neighbourhood_rating, n_persons))
```

After trying to get the predictions to work on the model generated using the ranger package, I eventually had to use the random forests package in order to get probabilities as I couldn't figure out how to get them to work with the ranger model.

```{r}
#create predictions object
predictions <- as.data.frame(predict(object = rf_nrating_classifier, test_neighbourhood_rating, type = "prob"))

#add probabilities to test dataset
test_neighbourhood_rating$pred <- predictions$Good

#Create roc object for random forest predictions on test dataset
roc_obj_nrating_tree <- test_neighbourhood_rating %>% 
  roc(response = neighbourhood_rating_binary, predictor = pred)#predictor is the probability column

#Plot roc object for random forest on test dataset
ggroc(data = roc_obj_nrating_tree, legacy.axes = TRUE)+
  labs(x = "False Positive Rate", y = "True positive rate")

#Get AUC for random forest performance on test dataset
auc(roc_obj_nrating_tree)
```


```{r}
#check the confusion table
test_nrating_pred %>% 
  tabyl(neighbourhood_rating_binary, pred) %>% 
  adorn_title()
```
```{r}
#true positive rate
10846/(10846 + 23)
#true negative rate
37/(37 + 761)
#false positive rate
761/(37 + 761)
#false negative rate
23/(23 + 10846)
```


## Logistic regression

```{r}
#modify dataset so it is appropriate for logistic regression
train_nrating_regression <- train_dataset_imputed %>% 
  #take out don't know b/c we want to predict good/poor ratings
  filter(neighbourhood_rating != "No opinion") %>% 
   #create binary version of neighbourhood rating
  mutate(neighbourhood_rating_binary = if_else(neighbourhood_rating %in% c(
    "Very good", "Fairly good"), "Good", "Poor"), .after = 
      neighbourhood_rating) %>%
  #factor all the categorical variables
  mutate(community_belonging = factor(community_belonging, levels = 
                                         c("Very strongly",
                                           "Fairly strongly",
                                           "Not very strongly",
                                           "Not at all strongly",
                                           "Don't know")),
         satisfaction_with_nearest_green_space = factor(
           satisfaction_with_nearest_green_space,
           levels = c("Very satisfied",
                      "Fairly satisfied",
                      "Neither satisfied nor dissatisfied",
                      "Fairly dissatisfied",
                      "Very dissatisfied",
                      "No opinion"
                      )),
         age = factor(age, levels = c(
           "16 - 34 Years",
           "35 - 64 Years",
           "65 + Years"
         )),
         gender = as.factor(gender),
         economic_status = as.factor(economic_status),
         highest_education_level = factor(highest_education_level, levels = 
                                            c(
                                              "Standard grade or equiv (SVQ level 1 or 2).",
                                              "Higher, A level or equivalent (SVQ Level 3)",
                                              "HNC/HND or equivalent (SVQ Level 4)",
                                              "Degree, Professional qualification (Above SVQ Level 4)",
                                              "Other qualification"
                                            )),
         nearest_green_space_use = as.factor(nearest_green_space_use),
         volunteering_last_twelve_months = as.factor(volunteering_last_twelve_months),
         neighbourhood_rating_binary = as.factor(neighbourhood_rating_binary)
         ) %>% 
  #change distance to nearest green space categories to match previous datasets
  mutate(distance_to_nearest_green_space = case_when(
    distance_to_nearest_green_space %in% c("A 5 minute walk or less",
                                           "Within a 6-10 minute walk") ~ "Within 10 minutes",
    distance_to_nearest_green_space %in% c("Within an 11-20 minute walk",
                                           "Within a 21-30 minute walk",
                                           "More than a 30 minute walk away") ~ "Over 10 minutes",
    TRUE ~ "Don't know"
  )) %>% 
  mutate(distance_to_nearest_green_space = factor(distance_to_nearest_green_space, levels = c(
    "Within 10 minutes",
    "Over 10 minutes",
    "Don't know"
  ))) %>% 
  #transform neighbourhood rating binary into logical variable
  mutate(neighbourhood_rating_good = neighbourhood_rating_binary == "Good") %>% 
  #take out original neighbourhood rating and the n_persons column (the number of people with the same observed characteristics isn't useful to the model)
  select(-c(neighbourhood_rating, n_persons, neighbourhood_rating_binary)) 
```

```{r}
#check for aliased variables in the dataset
alias(neighbourhood_rating_good ~ ., data = train_nrating_regression)
```


```{r}
#use glmulti to search for the best model
glmulti_search_all_mains <- glmulti(
  neighbourhood_rating_good ~ .,
  data = train_nrating_regression,
  level = 1,
  method = "h",
  crit = "bic",
  confsetsize = 10,
  plotty = F,
  report = T,
  fitfunction = "glm",
  family = binomial(link = "logit")
)

summary(glmulti_search_all_mains)
```

```{r}
top <- weightable(glmulti_search_all_mains)
```


```{r}
#assign best model to an object and get summary
mod_nrating_econ <- 
  glm(neighbourhood_rating_good ~ community_belonging + 
        satisfaction_with_nearest_green_space + economic_status + 
        highest_education_level + year + household_size,
      data = train_nrating_regression,
      family = binomial(link = "logit"))



mod_nrating_no_econ <- 
  glm(neighbourhood_rating_good ~ community_belonging + 
        satisfaction_with_nearest_green_space + 
        highest_education_level + year + household_size,
      data = train_nrating_regression,
      family = binomial(link = "logit"))

anova(mod_nrating_econ, mod_nrating_no_econ)

summary(mod_nrating_no_econ)
```
```{r}
#get odds ratio for community belonging = not at all strongly
exp(-3.16)
#Take this result and get a percentage
1-0.04242574
```
```{r}
stats::confint(mod_nrating_no_econ, 'household_size', level = 0.95)
```


### Get odds ratio for model coefficients

```{r}
#has function odds.ratio
library(questionr)
options(scipen = 1000)
odds.ratio(mod_nrating_no_econ, level = 0.95)
```





```{r}
#pull log likelihoods for null and proposed models
ll.null <- mod_nrating_no_econ$null.deviance/-2
ll.proposed <- mod_nrating_no_econ$deviance/-2
#get pseudo R^2 using this equation - this is the overall effect size
(ll.null - ll.proposed)/ll.null
```
```{r}
#get p value for R^2
1 - pchisq(2*(ll.proposed - ll.null), df=(length(mod_nrating_no_econ$coefficients)-1))
```


```{r}
#odds_ratio = exp(coefficient for categorical predictor)
#community belonging - not at all strongly
exp(3.100740)

```
The coefficient for community_belonging ratings of "not at all strongly" was -3.1. This corresponds to an odds ratio of -22.21. Therefore, the odds of selecting a "good" neighbourhood rating are 22.2% lower if the participant indicated that they belonged "not at all strongly" to their community when compared with participants who indicated they belonged "very strongly" to their community.

```{r}
train_nrating_regression %>% distinct(highest_education_level)
```



```{r}
exp(2.006943)
```
The coefficient for satisfaction with green space ratings of "very dissatisfied" was -2.0. This corresponds to an odds ratio of 7.44. Therefore, the odds of selecting a "good" neighbourhood rating are 7.4% lower if the participant indicated that they were very dissatisfied with their nearest green space when compared with participants who indicated they were very satisfied with their nearest green space.

```{r}
exp(0.662114)
```
```{r}
exp(0.064584)
```


### Confusion matrix
```{r}
nrating_regression_preds <- train_nrating_regression %>% 
  # add our predicted probabilities from the log model
  add_predictions(mod_nrating_no_econ, type = "response") %>% 
    mutate(predicted_good = pred >= 0.6)

conf_mat <- nrating_regression_preds %>% 
  tabyl(neighbourhood_rating_good, predicted_good) %>% 
  adorn_title()
```
```{r}
roc_obj_nrating <- nrating_regression_preds %>% 
  roc(response = neighbourhood_rating_good, predictor = pred)#predictor is the probability column

ggroc(data = roc_obj_nrating, legacy.axes = TRUE)+
  labs(x = "False Positive Rate", y = "True positive rate")

#get auc for training model
auc(roc_obj_nrating)
```
### Test on the test dataset

```{r}
#format the test dataset using the same method as the train dataset

test_nrating_regression <- test_dataset_imputed %>% 
  #take out don't know b/c we want to predict good/poor ratings
  filter(neighbourhood_rating != "No opinion") %>% 
   #create binary version of neighbourhood rating
  mutate(neighbourhood_rating_binary = if_else(neighbourhood_rating %in% c(
    "Very good", "Fairly good"), "Good", "Poor"), .after = 
      neighbourhood_rating) %>%
  #factor all the categorical variables
  mutate(community_belonging = factor(community_belonging, levels = 
                                         c("Very strongly",
                                           "Fairly strongly",
                                           "Not very strongly",
                                           "Not at all strongly",
                                           "Don't know")),
         satisfaction_with_nearest_green_space = factor(
           satisfaction_with_nearest_green_space,
           levels = c("Very satisfied",
                      "Fairly satisfied",
                      "Neither satisfied nor dissatisfied",
                      "Fairly dissatisfied",
                      "Very dissatisfied",
                      "No opinion"
                      )),
         age = factor(age, levels = c(
           "16 - 34 Years",
           "35 - 64 Years",
           "65 + Years"
         )),
         gender = as.factor(gender),
         economic_status = as.factor(economic_status),
         highest_education_level = factor(highest_education_level, levels = 
                                            c(
                                              "Standard grade or equiv (SVQ level 1 or 2).",
                                              "Higher, A level or equivalent (SVQ Level 3)",
                                              "HNC/HND or equivalent (SVQ Level 4)",
                                              "Degree, Professional qualification (Above SVQ Level 4)",
                                              "Other qualification"
                                            )),
         nearest_green_space_use = as.factor(nearest_green_space_use),
         volunteering_last_twelve_months = as.factor(volunteering_last_twelve_months),
         neighbourhood_rating_binary = as.factor(neighbourhood_rating_binary)
         ) %>% 
  #change distance to nearest green space categories to match previous datasets
  mutate(distance_to_nearest_green_space = case_when(
    distance_to_nearest_green_space %in% c("A 5 minute walk or less",
                                           "Within a 6-10 minute walk") ~ "Within 10 minutes",
    distance_to_nearest_green_space %in% c("Within an 11-20 minute walk",
                                           "Within a 21-30 minute walk",
                                           "More than a 30 minute walk away") ~ "Over 10 minutes",
    TRUE ~ "Don't know"
  )) %>% 
  mutate(distance_to_nearest_green_space = factor(distance_to_nearest_green_space, levels = c(
    "Within 10 minutes",
    "Over 10 minutes",
    "Don't know"
  ))) %>% 
  #transform neighbourhood rating binary into logical variable
  mutate(neighbourhood_rating_good = neighbourhood_rating_binary == "Good") %>% 
  #take out original neighbourhood rating and the n_persons column (the number of people with the same observed characteristics isn't useful to the model)
  select(-c(neighbourhood_rating, n_persons, neighbourhood_rating_binary)) 
```


```{r}
#add predictions to the test dataset using the model
test_nrating_regression_preds <- test_nrating_regression %>% 
   add_predictions(mod_nrating_no_econ, type = "response") %>% 
   mutate(predicted_good = pred >= 0.6)

#create roc object
roc_obj_nrating_regression_test <- test_nrating_regression_preds %>% 
  roc(response = neighbourhood_rating_good, predictor = pred)#predictor is the probability column

#plot the roc object
ggroc(data = roc_obj_nrating_regression_test, legacy.axes = TRUE)+
  labs(x = "False Positive Rate", y = "True positive rate")

#get auc for model performance on test data
auc(roc_obj_nrating_regression_test)
```





## Compare Logistic regression with Random forest

```{r}
#plot roc curves for both models on same graph
ggroc(list("Random forest" = roc_obj_nrating_tree, "Logistic regression" = roc_obj_nrating_regression_test))+
  labs(title = "True positive vs false positive rate for models",
       x = "False Positive Rate", y = "True positive rate") +
  scale_colour_manual(values = c("purple", "forest green"))

#Get AUC for regression on test dataset
auc(roc_obj_nrating_regression_test)

#Get AUC for random forest performance on test dataset
auc(roc_obj_nrating_tree)

#Get AUC confidence intervals for regression
ci.auc(roc_obj_nrating_regression_test)

#Get AUC confidence intervals for random forest
ci.auc(roc_obj_nrating_tree)
```

