---
title: "R Notebook"
output: html_notebook
---
# Libraries
```{r}
library(tidyverse)
library(janitor)
library(missRanger)
library(ranger)
library(dlookr)

```
#Data
```{r}
aggregate_dataset <- read_csv("data/shs_aggregate_responses.csv") %>% clean_names()
```


## Notes about the dataset
- There are some NAs in the data, namely in the following rows
    - Satisfaction with nearest greenspace (9988 missing)
    - Economic status (1 missing)
    - Highest education level (10920 missing)
    - Nearest green space use (16766 missing)
    - Volunteering last twelve months (6778)
    
Can we meaningfully impute these?

- 
```{r}
plot_na_pareto(aggregate_dataset)
```
```{r}
plot_na_intersect(aggregate_dataset)
```


```{r}
aggregate_dataset_imputed <- aggregate_dataset %>% 
  mutate(volunteering_last_twelve_months = 
           if_else(is.na(volunteering_last_twelve_months), 
                   "Don't know", volunteering_last_twelve_months),
         nearest_green_space_use = 
           if_else(is.na(nearest_green_space_use), 
                   "Don't know", nearest_green_space_use),
         highest_education_level = 
           if_else(is.na(highest_education_level), "Don't know",
                   highest_education_level),
         satisfaction_with_nearest_green_space = 
           if_else(is.na(satisfaction_with_nearest_green_space), "No opinion",
                   satisfaction_with_nearest_green_space)
         ) %>% 
  drop_na()
```

```{r}
aggregate_dataset_imputed %>% 
  filter(community_belonging != "Don't know") %>% 
  mutate(community_belonging_binary = if_else(community_belonging %in% c(
    "Very strongly", "Fairly strongly"), "Strong", "Not strong"), .after = 
      community_belonging) 
```

Community belonging - Don't know - 576
Neighbourhood rating - No opinion - 190
Satisfaction with nearest greenspace - No opinion - 3418
Nearest green space use - Don't know - 83
Distance to nearest green space - Don't know - 386


To Impute or not Impute: That’s the Question - Lodder, 2013
"If more than
25% of the data is missing and researchers apply modern
treatments to impute the missing data, then they should
always compare the results of their subsequent analyses
with the results they would have obtained if they had
used complete case analysis." (Lodder, 2013)


So we can try using the MissRanger package to compute missing categorical variables using the random forest method (https://yuzar-blog.netlify.app/posts/2021-03-04-how-to-impute-missing-values-in-r/#multivariate-imputation-of-categorical-variables)

We can then compare the results to the complete case analysis (only complete observations are included - meaning all NAs are dropped)

## Test/train split

```{r}
n_data <- nrow(aggregate_dataset)
test_index <- sample(1:n_data, size = n_data*0.2)
agg_dataset_test <- slice(aggregate_dataset, test_index)
agg_dataset_train <- slice(aggregate_dataset, -test_index)
```

### Imputation for train dataset
```{r}
train_dataset_imputed <- missRanger(
  agg_dataset_train,
  formula = . ~ .,
  num.trees = 1000,
  verbose = 2,
  seed = 111,
  returnOOB = TRUE
  
)
```
### Imputation for test dataset
```{r}
test_dataset_imputed <- missRanger(
  agg_dataset_test,
  formula = . ~ .,
  num.trees = 1000,
  verbose = 2,
  seed = 111,
  returnOOB = TRUE
)
```




```{r}
train_comm_belonging <- train_dataset_imputed %>% 
  #take out don't know b/c we want to predict strong/weak belonging
  filter(community_belonging != "Don't know") %>% 
   #create binary version of community belonging
  mutate(community_belonging_binary = if_else(community_belonging %in% c(
    "Very strongly", "Fairly strongly"), "Strong", "Not strong"), .after = 
      community_belonging) %>%
  #factor all the categorical variables
  mutate(neighbourhood_rating = factor(neighbourhood_rating, levels = 
                                         c("Very good",
                                           "Fairly good",
                                           "Fairly poor",
                                           "Very poor",
                                           "No opinion")),
         distance_to_nearest_green_space = factor(distance_to_nearest_green_space,
                                                  levels = c(
                                                    "A 5 minute walk or less",
                                                    "Within a 6-10 minute walk",
                                                    "Within an 11-20 minute walk",
                                                    "Within a 21-30 minute walk",
                                                    "More than a 30 minute walk away",
                                                    "Don't know"
                                                  )),
         satisfaction_with_nearest_green_space = factor(
           satisfaction_with_nearest_green_space,
           levels = c("Very satisfied",
                      "Fairly satisfied",
                      "Neither satisfied nor dissatisfied",
                      "Fairly dissatisfied",
                      "Very dissatisfied",
                      "No opinion"
                      )),
         age = factor(age, levels = c(
           "16 - 34 Years",
           "35 - 64 Years",
           "65 + Years"
         )),
         gender = as.factor(gender),
         economic_status = as.factor(economic_status),
         highest_education_level = factor(highest_education_level, levels = 
                                            c(
                                              "Standard grade or equiv (SVQ level 1 or 2).",
                                              "Higher, A level or equivalent (SVQ Level 3)",
                                              "HNC/HND or equivalent (SVQ Level 4)",
                                              "Degree, Professional qualification (Above SVQ Level 4)",
                                              "Other qualification"
                                            )),
         nearest_green_space_use = as.factor(nearest_green_space_use),
         volunteering_last_twelve_months = as.factor(volunteering_last_twelve_months),
         community_belonging_binary = as.factor(community_belonging_binary)
         ) %>% 
  #take out original community belonging and the n_persons column (the number of people with the same observed characteristics isn't useful to the model)
  select(-c(community_belonging, n_persons))
```

```{r}
aggregate_dataset_imputed %>% distinct(nearest_green_space_use)
```

```{r}
community_rf_classifier <- ranger(community_belonging_binary ~ .,
                                  data = train_comm_belonging,
                                  importance = "impurity",
                                  num.trees = 1000)
```

```{r}
community_rf_classifier
```

```{r}
importance(community_rf_classifier)
```

```{r}
community_rf_classifier$confusion.matrix
```

```{r}
#testing to see whether a different value for mtry will be better than the default
oob_values <- vector(length = 10)
for(i in 1:10) {
  temp.model <- ranger(community_belonging_binary ~ .,
                       data = train_comm_belonging,
                       importance = "impurity",
                       num.trees = 1000,
                       mtry = i)
  oob_values[i] <- temp.model$prediction.error
}
oob_values
```

