---
title: "R Notebook"
output: html_notebook
---
# Libraries
```{r}
library(tidyverse)
library(janitor)
```
#Data
```{r}
aggregate_dataset <- read_csv("data/shs_aggregate_responses.csv") %>% clean_names()
```

```{r}
aggregate_dataset %>% head()
```

```{r}
aggregate_dataset %>% skimr::skim()
```

```{r}
aggregate_dataset %>% distinct(volunteering_last_twelve_months)
```
## Notes about the dataset
- There are some NAs in the data, namely in the following rows
    - Satisfaction with nearest greenspace (9988 missing)
    - Economic status (1 missing)
    - Highest education level (10920 missing)
    - Nearest green space use (16766 missing)
    - Volunteering last twelve months (6778)
    
Can we meaningfully impute these?

- 

```{r}
aggregate_dataset %>% 
  filter(is.na(highest_education_level)) %>% 
  count(economic_status)
```


```{r}
aggregate_dataset_imputed <- aggregate_dataset %>% 
  mutate(volunteering_last_twelve_months = 
           if_else(is.na(volunteering_last_twelve_months), 
                   "Don't know", volunteering_last_twelve_months),
         nearest_green_space_use = 
           if_else(is.na(nearest_green_space_use), 
                   "Don't know", nearest_green_space_use),
         highest_education_level = 
           if_else(is.na(highest_education_level), "Don't know",
                   highest_education_level),
         satisfaction_with_nearest_green_space = 
           if_else(is.na(satisfaction_with_nearest_green_space), "No opinion",
                   satisfaction_with_nearest_green_space)
         ) %>% 
  drop_na()
```

```{r}
aggregate_dataset_imputed %>% 
  filter(community_belonging != "Don't know") %>% 
  mutate(community_belonging_binary = if_else(community_belonging %in% c(
    "Very strongly", "Fairly strongly"), "Strong", "Not strong"), .after = 
      community_belonging) 
```

Community belonging - Don't know - 576
Neighbourhood rating - No opinion - 190
Satisfaction with nearest greenspace - No opinion - 3418
Nearest green space use - Don't know - 83
Distance to nearest green space - Don't know - 386